---
title: 'Statistical report: EF reclassification in HFmrEF and 5% digit bias across EF categories'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2: 
    dev: cairo_pdf
    fig_caption: yes
    fig_height: 7
    fig_width: 8
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
urlcolor: blue
linkcolor: black
header-includes:
   - \usepackage{subfig}
   - \usepackage{pdflscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
---

\clearpage
\listoftables
\listoffigures
\clearpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

hf3_lt_fu_data_soladis_jan19.sas7bdat (from the folder DATABASE_ANALYZED.ZIP). 

## Inclusion/exclusion criteria

```{r flow}
default_kable(flow, caption = "Flowchart", scale_down = F)
```

First patient in: `r min(pdata$num_dmVisitdt)` and last patient in: `r max(pdata$num_dmVisitdt)`. 

The median age (IQR) is `r pdata %>% summarise(med = fn(median(num_age, na.rm = T), dig = 1),
                                             q1 = fn(quantile(num_age, probs = 0.25, na.rm = T), dig = 1),
                                             q3 = fn(quantile(num_age, probs = 0.75, na.rm = T), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r pdata %>% count(num_dmgender) %>%
  mutate(perc = fn(n / sum(n) * 100, 1)) %>%
  filter(num_dmgender == "Female") %>%
  pull(perc)`% females.  
  
## General

In those cases variables have been collected at different time points the in-hospital (not discharge) 
measure is used for hospitalized patients. 

If not stated overwise EF is EF from Echo-Doppler performed during the hospitalization/visit.

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

\clearpage

## EF overview

```{r, child = "./src/tabefflow.Rmd"}

```

\clearpage

## Baseline characteristics

```{r, child = "./src/tab1.Rmd"}

```

\clearpage

## Baseline characteristics 5% digit

```{r, child = "./src/tab05digit.Rmd"}

```

\clearpage

```{r, child = "./src/bar.Rmd"}

```

\clearpage

### "Bias"

In order to quantify the bias of EF reported with 0 or 5 as last digits we assumed 
that EF followed 1) a gamma distribution with shape 7.4 and rate 0.2 
2) a normal distribution with mean 38.6 and standard deviation 13.9. 
The parameters were based on the maximum-likelihood fitting of the data 
assuming the data followed a normal distribution. A kernel density estimate with a 
smoothing bandwidth of 2 was also computed. The density estimates were 
selected for values between 0 and 100. 
The difference between the observed percentages in each EF 1% increment and what 
would have been expected given that EF had been recorded evenly across 
the spectrum was thereafter calculated and summarised. 

```{r, child = "./src/bardensity.Rmd"}

```

\clearpage

## Outcome analysis

The following outcomes are considered: 

- All-cause mortality
- Cardiovascular mortality
- Cardiovascular mortality/First HF hospitalization
- First HF hospitalization

Incidence per 1000 py was calculated with 95% Poisson confidence intervals. 

All outcome analysis are performed excluding patients that were lost to follow-up. 

For patients with missing information on the date of hospitalisation 
the time to hospitalisation was imputed with half the time to last
follow-up. Further, times to hospitalisation that were larger than time 
to death or last follow-up were set to death or last follow-up. 

Patients with missing information on cause of death or hospitalisation 
are not included in the CV death/HF hospitalisation. 

Time is from date of discharge for hospitalized patients, or if this is missing, date of admission and 
censored at death not defined as an event or follow-up.  

The outcomes are presented with the cumulative incidence curves with HR (95% CI) from crude cox proportional hazards regression.

The median (min-max) follow-up is `r pdata %>% filter(survpop) %>% summarise(med = fn(median(outtime_death / 365.25 * 12), dig = 1),
                                             min = fn(min(outtime_death / 365.25 * 12), dig = 1),
                                             max = fn(max(outtime_death / 365.25 * 12), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` months for a total of 
                                   `r pdata %>% filter(survpop) %>% summarise(sumpy = fn(sum(outtime_death) / 365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.


```{r, child = "./src/incidencetab.Rmd"}

```

```{r, child = "./src/km.Rmd"}

```

\clearpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/esc_efreclass. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
